### 2016-07-15 testbed_http.tex
provided more text about http method

\begin{lstlisting}[]
$ export HTTPBOOTIMAGE="${WORKDIR}/httpboot"
$ export HTTPROOTIMAGE="${WORKDIR}/httproot"
$ export HTTPHOMEIMAGE="${WORKDIR}/httphome"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

The initial \ac{RAM} filesystem contains an \texttt{init} script that
sets up the prepares the yis first
called. This script initializes the  may be rewritten, but

Create the following file (THE READ-ONLY OPTION IS STILL NOT CONFIGURABLE IN THIS SCRIPT)
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/etc/initramfs-tools/scripts/http}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
# Fetch, cache, and mount root image using HTTP

@@

    # Mount root as readonly
    losetup /dev/loop0 $(readlink -f ${IMAGEPATH})
    mount -r -n -t ext4 -o nodiratime,noatime /dev/loop0 /root

    # Bind /root/home to /home
    #mount -n -o rbind /home /root/home
}

@@

%Add the below configurations to \$\{HTTPROOTDIR\}/boot/cmdline.txt
(remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/boot/cmdline.txt}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
<@\Suppressnumber @>boot=http ip=dhcp image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0
    url=http://kom.aau.dk/project/TuneSCode/raspi/<@\Reactivatenumber @>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% ip=dhcp root=/dev/http http-options=hard,intr,ro

Here is how the variables are used in the initramfs:
\begin{itemize}
    \item boot=http: tells the init script in initramfs to call the \texttt{http} script presented above
    \item ip=dhcp: use \ac{DHCP} to obtain an \ac{IP} address in the initial ram filesystem
    \item image=<device>:<root image path>: is where the \ac{Raspi} will store the downloaded root image.
    \item root=/dev/loop0: The \texttt{http} initramfs script will make the root filesystem available on /dev/loop0
    \item url=<URL to root image>: Is the directory on a \ac{HTTP} server where the root image will be iavailable.
\end{itemize}

>>>>

\begin{lstlisting}[]
$ export HTTPBOOTIMAGE="${WORKDIR}/httpboot"
$ export HTTPROOTIMAGE="${WORKDIR}/httproot"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

The initial \ac{RAM} filesystem contains an \texttt{init} script that
is first called to initializes
\texttt{initramfs}, parse arguments specified in \texttt{cmdline.txt}, and
call a specialized script that prepares the actual root filesystem. By default,
Raspbian includes scripts to mount the root filesystem from 1) a block devices
(i.e. the memory card) and 2) from a \ac{NFS} shared folder. We make a third
script to mount the root filesystem from an image
(i.e. the root image \texttt{httproot.img}) that is downloaded from a
\ac{HTTP} server.

The initramfs script to download and mount the image from a \ac{HTTP} server
will require a few additional arguments. E.g. to specify from which \ac{HTTP}
server to download the root filesystem image from.
%This script that we will create will require a few additional arguments such
%as which \ac{HTTP} server the image should be downloaded from.
These arguments
may be specified in \texttt{cmdline.txt}.
%and later parsed within the \texttt{initramfs} script.

%Add the below configurations to \$\{HTTPROOTDIR\}/boot/cmdline.txt
\textcolor{red}{(remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)}
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/boot/cmdline.txt}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
<@\Suppressnumber @>boot=http ip=dhcp image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0
    url=http://kom.aau.dk/project/TuneSCode/raspi/<@\Reactivatenumber @>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% ip=dhcp root=/dev/http http-options=hard,intr,ro

Here is how the variables will be used in the \texttt{initramfs}:
\begin{itemize}
    \item boot=http: tells the init script in initramfs to call the \texttt{http} script that will be presented below
    \item ip=dhcp: use \ac{DHCP} to obtain an \ac{IP} address in the \texttt{initramfs}
    \item image=<device>:<root image path>: is where the \ac{Raspi} will store the downloaded root image.
    \item root=/dev/loop0: The \texttt{http} initramfs script will make the root filesystem available on /dev/loop0
    \item url=<URL to root image>: Is the directory on a \ac{HTTP} server where the root image will be iavailable.
\end{itemize}

The specialized script that we made to download and mount the root filesystem image from a \ac{HTTP} server should be created in the root filesystem as presented below:
%Create the following file
\textcolor{red}{(THE READ-ONLY OPTION IS STILL NOT CONFIGURABLE IN THIS SCRIPT. I'M NOT EVEN SURE IT IS POSSIBLE TO MAKE THE IMAGE READ-WRITABLE)}
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/etc/initramfs-tools/scripts/http}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
# Fetch, cache, and mount root image using HTTP

@@

    # Mount root as readonly
    losetup /dev/loop0 $(readlink -f ${IMAGEPATH})
    mount -r -n -t ext4 -o nodiratime,noatime /dev/loop0 /root
}

@@

The variable (boot=http) that we specified in \texttt{cmdline.txt} tells the
\texttt{init} script to call the script above rather than the default scripts
that mounts the root filesystem from a block device or a \ac{NFS} shared
folder. The \texttt{init} script will call the function \texttt{http\_mount\_root}
which in this case can be consired as the \texttt{main} function.

As in any other shell scripts, this means that code in the global namespace
will be executed by the interpretor before the function \texttt{http\_mount\_root}
is entered.
%Due to the variable that we specified in \texttt{cmdline.txt}, 
Hence, line 9 will be the first command executed. This line creates a directory
in \texttt{initramfs} named
\texttt{imagedir} that we will use to hold the root image.

Next, line 12 to line 44 will parse the additional arguments that we provided in
\texttt{cmdline.txt}.

The \ac{Raspi}s does not have enough \ac{RAM} to hold entire root
filesystem, so we instead download it to the partition that, as specified in
\texttt{fstab} later, will later be mounted as the \texttt{HOME} directory
when the system starts. This will enable the \ac{Raspi}s to cache the root image
and use the previously downloaded root image again after reboots (see the
function \texttt{http\_fetch\_image} starting from line 56)\fxnote{is there a better way to reference to a function within the script? This reference will be invalid if the script is slightly modified}.
